---
title: "BSE658A-Assignment-03"
author: "Abin Thomas (Roll No. 21218261)"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
library(tidyverse)
```

### Questions

1.  If you open the R console, the first line in the console shows the version of R you have installed on your computer. Please specify which R version you have by opening R console. (1 point)

Answer: *4.2.1*

2.  There are 'base' packages in R which you don't have to install separately after installing the latest version of R - they are bundled with R (like the furniture or dining table you might get when you rent a new house.) One of these base packages/libraries in R is `datasets`. Type `help(datasets)` to read more about it. View the complete list of datasets in `datasets` and load the dataset `AirPassengers` in your RStudio. Write the code below which you will use to solve question 3 and 4. (1 point)

Answer:

```{r}
AirPassengers
help("AirPassengers")
```

3.  Mention the source of the dataset `AirPassengers`. (1 point)

Answer: *Box, G. E. P., Jenkins, G. M. and Reinsel, G. C. (1976) Time Series Analysis, Forecasting and Control. Third Edition. Holden-Day. Series G*

4.  How many rows and columns does the dataset `AirPassengers` have? (1 point)

Answer: Number of rows: *12*, number of columns: *12*.

```{r}
apmatrix <- matrix(AirPassengers, ncol = 12, byrow =TRUE,
dimnames = list( as.character(1949:1960),month.abb))
apmatrix
```

```{r}
aptibble<- as_tibble(apmatrix)
nrow(apmatrix)
```

```{r}
ncol(apmatrix)
```

5.  There are multiple columns in `AirPassengers`. Can you create a new column `Passengers` by adding all the values in different columns in a given row? The number of rows in the original dataset and the new dataset will be the same. (3 points)

Answer:

```{r}
apt1 <- aptibble %>% mutate(Passengers = rowSums(.),.before=Jan)
apt1
```

6.  How will you check the number of `NA` values in the dataset `AirPassengers`? (1 point)

Answer:

```{r}
sum(is.na(apt1))
```

7.  You might have used `Tidyverse` package to solve the above questions. Can you create a new column `PercentPassengers` which shows the percentage of passengers who flew in a given month of a given year? The value of hundred percentage denotes the total number of passengers who flew in a given year (which you have already calculated in the column `Passengers`). Make sure the column `PercentPassengers` is the first column in the dataset. (3 points)

Answer:

```{r}
apt2 <- apt1 %>% mutate(across()*100/apt1$Passengers )
Pss=apt1$Passengers
#stored the list of passengers across years in a list 'Pss'
#because haven't discovered how to use 'across' with exceptions
apt2 <- apt2 %>% mutate(Passengers = Pss)
#plugged the number of passengers for each year back into the final tibble
apt2
```
